---
title: 'Q2: Does climate drive lake-landscape coherence?'
author: "Jonathan Walter, Grace Wilkinson, Rachel Fleck, Michael Pace"
date: "12/9/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(aqts)
library(wsyn)
library(rgdal)
library(rgeos)
library(LAGOSNE)
library(raster)
library(dplyr)
library(viridis)

rm(list=ls())

source("/Users/jonathanwalter/GitHub/AquaTerrSynch/AnalysisCode/Q2_ClimateCoherence/bandtest_coh.R")

```

This document organizes for openness and reproducibility analyses of the influence of climate .

# Data import

In the Q1 manuscript, we output time series from coherent lake-landscape pairs. These are loaded here. We also load time series of three major climate indices (ENSO, PDO, NAO) and local temperature and precipitation.

```{r load and format data, echo=TRUE, cache=TRUE}

#load coherent lakes data
load("~/Box Sync/NSF EAGER Synchrony/Data/RData files/q2_coherent_lakes.RData")
#load climate index time series
mei.raw<-read.csv("/Users/jonathanwalter/Documents/Research/DATA/ClimateIndices/mei_monthly_1979_2019.csv")#MEI
pdo.raw<-read.csv("/Users/jonathanwalter/Documents/Research/DATA/ClimateIndices/pdo_monthly_1950_2019.csv")#PDO
nao.raw<-read.csv("/Users/jonathanwalter/Documents/Research/DATA/ClimateIndices/nao_monthly_1950_2019.csv")#NAO

mei.gs<-mei.raw[mei.raw$mm >=5 & mei.raw$mm <=9,]
pdo.gs<-pdo.raw[pdo.raw$mm >=5 & pdo.raw$mm <=9,]
nao.gs<-nao.raw[nao.raw$mm >=5 & nao.raw$mm <=9,]

mei.gs<-aggregate(mei.gs$mei,list(mei.gs$yyyy),FUN="mean")
names(mei.gs)<-c("year","mei")
pdo.gs<-aggregate(pdo.gs$PDO,list(pdo.gs$yyyy),FUN="mean")
names(pdo.gs)<-c("year","pdo")
nao.gs<-aggregate(nao.gs$NAO,list(nao.gs$yyyy),FUN="mean")
names(nao.gs)<-c("year","nao")

for(ii in 1:length(coherentlakes.st$lakedata)){
  coherentlakes.st$lakedata[[ii]]<-rbind(coherentlakes.st$lakedata[[ii]],
                                         mei.gs$mei[mei.gs$year %in% colnames(coherentlakes.st$lakedata[[ii]])],
                                         pdo.gs$pdo[pdo.gs$year %in% colnames(coherentlakes.st$lakedata[[ii]])],
                                         nao.gs$nao[nao.gs$year %in% colnames(coherentlakes.st$lakedata[[ii]])])
  rownames(coherentlakes.st$lakedata[[ii]])<-c("chla","ndvi","mei","pdo","nao")
}

for(ii in 1:length(coherentlakes.lt$lakedata)){
  coherentlakes.lt$lakedata[[ii]]<-rbind(coherentlakes.lt$lakedata[[ii]],
                                         mei.gs$mei[mei.gs$year %in% colnames(coherentlakes.lt$lakedata[[ii]])],
                                         pdo.gs$pdo[pdo.gs$year %in% colnames(coherentlakes.lt$lakedata[[ii]])],
                                         nao.gs$nao[nao.gs$year %in% colnames(coherentlakes.lt$lakedata[[ii]])])
  rownames(coherentlakes.lt$lakedata[[ii]])<-c("chla","ndvi","mei","pdo","nao")
}

#add local weather
ppt.files<-paste0("/Users/jonathanwalter/Documents/Research/DATA/PRISM_4km2_gridded/ppt/PRISM_ppt_stable_4kmM3_",
                  rep(1989:2018,each=5),
                  rep(c("05","06","07","08","09"),times=30),
                  "_bil.bil")
ppt<-stack(ppt.files)


tavg.files<-c(paste0("/Users/jonathanwalter/Documents/Research/DATA/PRISM_4km2_gridded/tmean/PRISM_tmean_stable_4kmM2_",
                  rep(1989:2016,each=5),
                  rep(c("05","06","07","08","09"),times=28),
                  "_bil.bil"),
              paste0("/Users/jonathanwalter/Documents/Research/DATA/PRISM_4km2_gridded/tmean/PRISM_tmean_stable_4kmM3_",
                  rep(2017:2018,each=5),
                  rep(c("05","06","07","08","09"),times=2),
                  "_bil.bil"))
tavg<-stack(tavg.files)

coherentlakes.st<-addPRISMts(coherentlakes.st, ppt, var="ppt")
coherentlakes.st<-addPRISMts(coherentlakes.st, tavg, var="tmean")
coherentlakes.lt<-addPRISMts(coherentlakes.lt, ppt, var="ppt")
coherentlakes.lt<-addPRISMts(coherentlakes.lt, tavg, var="tmean")
```
Make some maps to see where we are
```{r mapping, echo=T, cache=T}
states<-readOGR("~/Box Sync/NSF EAGER Synchrony/Data/statesp020_nolakes.shp")
getstates<-c("Minnesota", "Iowa", "Wisconsin", "Illinois", "Missouri", "Michigan", "Indiana", "Ohio", "Pennsylvania", "New York", "New Jersey", "Connecticut", "New Hampshire", "Rhode Island", "Massachusetts", "Vermont", "Maine")
lagosstates<-states[states@data$STATE %in% getstates,]

plot(lagosstates, main="Coherent lakes, short timescales")
points(coherentlakes.st$lakeinfo$nhd_long, coherentlakes.st$lakeinfo$nhd_lat, pch=16, col="blue")

plot(lagosstates, main="Coherent lakes, long timescales")
points(coherentlakes.lt$lakeinfo$nhd_long, coherentlakes.lt$lakeinfo$nhd_lat, pch=16, col="blue")


```

# Analyses

a) Are lakes and landscapes coherent with the same climate variables? b) Do climate variables explain a similar amount of variability in coherent lakes as the landscape does? c) does this differ by timescale?

First, do coherences with climate variables
```{r climate coherence, echo=TRUE, cache=FALSE}
lakeXclim.st<-data.frame(lagoslakeid=names(coherentlakes.st$lakedata),
                         mei.coh=rep(NA, length(coherentlakes.st$lakedata)),
                         mei.p=rep(NA, length(coherentlakes.st$lakedata)),
                         mei.phi=rep(NA, length(coherentlakes.st$lakedata)),
                         nao.coh=rep(NA, length(coherentlakes.st$lakedata)),
                         nao.p=rep(NA, length(coherentlakes.st$lakedata)),
                         nao.phi=rep(NA, length(coherentlakes.st$lakedata)),
                         pdo.coh=rep(NA, length(coherentlakes.st$lakedata)),
                         pdo.p=rep(NA, length(coherentlakes.st$lakedata)),
                         pdo.phi=rep(NA, length(coherentlakes.st$lakedata)),
                         ppt.coh=rep(NA, length(coherentlakes.st$lakedata)),
                         ppt.p=rep(NA, length(coherentlakes.st$lakedata)),
                         ppt.phi=rep(NA, length(coherentlakes.st$lakedata)),
                         tavg.coh=rep(NA, length(coherentlakes.st$lakedata)),
                         tavg.p=rep(NA, length(coherentlakes.st$lakedata)),
                         tavg.phi=rep(NA, length(coherentlakes.st$lakedata)))
landXclim.st<-lakeXclim.st

lakeXclim.lt<-data.frame(lagoslakeid=names(coherentlakes.lt$lakedata),
                         mei.coh=rep(NA, length(coherentlakes.lt$lakedata)),
                         mei.p=rep(NA, length(coherentlakes.lt$lakedata)),
                         mei.phi=rep(NA, length(coherentlakes.lt$lakedata)),
                         nao.coh=rep(NA, length(coherentlakes.lt$lakedata)),
                         nao.p=rep(NA, length(coherentlakes.lt$lakedata)),
                         nao.phi=rep(NA, length(coherentlakes.lt$lakedata)),
                         pdo.coh=rep(NA, length(coherentlakes.lt$lakedata)),
                         pdo.p=rep(NA, length(coherentlakes.lt$lakedata)),
                         pdo.phi=rep(NA, length(coherentlakes.lt$lakedata)),
                         ppt.coh=rep(NA, length(coherentlakes.lt$lakedata)),
                         ppt.p=rep(NA, length(coherentlakes.lt$lakedata)),
                         ppt.phi=rep(NA, length(coherentlakes.lt$lakedata)),
                         tavg.coh=rep(NA, length(coherentlakes.lt$lakedata)),
                         tavg.p=rep(NA, length(coherentlakes.lt$lakedata)),
                         tavg.phi=rep(NA, length(coherentlakes.lt$lakedata)))
landXclim.lt<-lakeXclim.lt

# lakes, short timescales
NN=length(coherentlakes.st$lakedata)
ts=c(2,4)
for(ii in 1:NN){
  vars<-rownames(coherentlakes.st$lakedata[[ii]])
  yy<-as.numeric(colnames(coherentlakes.st$lakedata[[ii]]))
  clnd<-cleandat(coherentlakes.st$lakedata[[ii]],yy,clev=5)$cdat

  cohXmei<-coh(clnd[vars=="chla",], clnd[vars=="mei"], yy, norm="powall", sigmethod="fast")
  cohXnao<-coh(clnd[vars=="chla",], clnd[vars=="nao"], yy, norm="powall", sigmethod="fast")
  cohXpdo<-coh(clnd[vars=="chla",], clnd[vars=="pdo"], yy, norm="powall", sigmethod="fast")
  cohXppt<-coh(clnd[vars=="chla",], clnd[vars=="ppt"], yy, norm="powall", sigmethod="fast")
  cohXtavg<-coh(clnd[vars=="chla",], clnd[vars=="tavg"], yy, norm="powall", sigmethod="fast")
  
  cohXmei<-bandtest.coh(cohXmei,ts)
  cohXnao<-bandtest.coh(cohXnao,ts)
  cohXpdo<-bandtest.coh(cohXpdo,ts)
  cohXppt<-bandtest.coh(cohXppt,ts)
  cohXtavg<-bandtest.coh(cohXtavg,ts)
  
  lakeXclim.st$mei.coh[ii]<-cohXmei$bandp$mn_coh
  lakeXclim.st$mei.p[ii]<-cohXmei$bandp$p_val
  lakeXclim.st$mei.phi[ii]<-cohXmei$bandp$mn_phs
  lakeXclim.st$nao.coh[ii]<-cohXnao$bandp$mn_coh
  lakeXclim.st$nao.p[ii]<-cohXnao$bandp$p_val
  lakeXclim.st$nao.phi[ii]<-cohXnao$bandp$mn_phs
  lakeXclim.st$pdo.coh[ii]<-cohXpdo$bandp$mn_coh
  lakeXclim.st$pdo.p[ii]<-cohXpdo$bandp$p_val
  lakeXclim.st$pdo.phi[ii]<-cohXpdo$bandp$mn_phs
  lakeXclim.st$ppt.coh[ii]<-cohXppt$bandp$mn_coh
  lakeXclim.st$ppt.p[ii]<-cohXppt$bandp$p_val
  lakeXclim.st$ppt.phi[ii]<-cohXppt$bandp$mn_phs
  lakeXclim.st$tavg.coh[ii]<-cohXtavg$bandp$mn_coh
  lakeXclim.st$tavg.p[ii]<-cohXtavg$bandp$p_val
  lakeXclim.st$tavg.phi[ii]<-cohXtavg$bandp$mn_phs
}

#land, short timescales
NN=length(coherentlakes.st$lakedata)
ts=c(2,4)
for(ii in 1:NN){
  vars<-rownames(coherentlakes.st$lakedata[[ii]])
  yy<-as.numeric(colnames(coherentlakes.st$lakedata[[ii]]))
  clnd<-cleandat(coherentlakes.st$lakedata[[ii]],yy,clev=5)$cdat

  cohXmei<-coh(clnd[vars=="ndvi",], clnd[vars=="mei"], yy, norm="powall", sigmethod="fast")
  cohXnao<-coh(clnd[vars=="ndvi",], clnd[vars=="nao"], yy, norm="powall", sigmethod="fast")
  cohXpdo<-coh(clnd[vars=="ndvi",], clnd[vars=="pdo"], yy, norm="powall", sigmethod="fast")
  cohXppt<-coh(clnd[vars=="ndvi",], clnd[vars=="ppt"], yy, norm="powall", sigmethod="fast")
  cohXtavg<-coh(clnd[vars=="ndvi",], clnd[vars=="tavg"], yy, norm="powall", sigmethod="fast")
  
  cohXmei<-bandtest.coh(cohXmei,ts)
  cohXnao<-bandtest.coh(cohXnao,ts)
  cohXpdo<-bandtest.coh(cohXpdo,ts)
  cohXppt<-bandtest.coh(cohXppt,ts)
  cohXtavg<-bandtest.coh(cohXtavg,ts)
  
  landXclim.st$mei.coh[ii]<-cohXmei$bandp$mn_coh
  landXclim.st$mei.p[ii]<-cohXmei$bandp$p_val
  landXclim.st$mei.phi[ii]<-cohXmei$bandp$mn_phs
  landXclim.st$nao.coh[ii]<-cohXnao$bandp$mn_coh
  landXclim.st$nao.p[ii]<-cohXnao$bandp$p_val
  landXclim.st$nao.phi[ii]<-cohXnao$bandp$mn_phs
  landXclim.st$pdo.coh[ii]<-cohXpdo$bandp$mn_coh
  landXclim.st$pdo.p[ii]<-cohXpdo$bandp$p_val
  landXclim.st$pdo.phi[ii]<-cohXpdo$bandp$mn_phs
  landXclim.st$ppt.coh[ii]<-cohXppt$bandp$mn_coh
  landXclim.st$ppt.p[ii]<-cohXppt$bandp$p_val
  landXclim.st$ppt.phi[ii]<-cohXppt$bandp$mn_phs
  landXclim.st$tavg.coh[ii]<-cohXtavg$bandp$mn_coh
  landXclim.st$tavg.p[ii]<-cohXtavg$bandp$p_val
  landXclim.st$tavg.phi[ii]<-cohXtavg$bandp$mn_phs
}

#lakes, long timescales
NN=length(coherentlakes.lt$lakedata)
ts=c(4,Inf)
for(ii in 1:NN){
  vars<-rownames(coherentlakes.lt$lakedata[[ii]])
  yy<-as.numeric(colnames(coherentlakes.lt$lakedata[[ii]]))
  clnd<-cleandat(coherentlakes.lt$lakedata[[ii]],yy,clev=5)$cdat

  cohXmei<-coh(clnd[vars=="chla",], clnd[vars=="mei"], yy, norm="powall", sigmethod="fast")
  cohXnao<-coh(clnd[vars=="chla",], clnd[vars=="nao"], yy, norm="powall", sigmethod="fast")
  cohXpdo<-coh(clnd[vars=="chla",], clnd[vars=="pdo"], yy, norm="powall", sigmethod="fast")
  cohXppt<-coh(clnd[vars=="chla",], clnd[vars=="ppt"], yy, norm="powall", sigmethod="fast")
  cohXtavg<-coh(clnd[vars=="chla",], clnd[vars=="tavg"], yy, norm="powall", sigmethod="fast")
  
  cohXmei<-bandtest.coh(cohXmei,ts)
  cohXnao<-bandtest.coh(cohXnao,ts)
  cohXpdo<-bandtest.coh(cohXpdo,ts)
  cohXppt<-bandtest.coh(cohXppt,ts)
  cohXtavg<-bandtest.coh(cohXtavg,ts)
  
  lakeXclim.lt$mei.coh[ii]<-cohXmei$bandp$mn_coh
  lakeXclim.lt$mei.p[ii]<-cohXmei$bandp$p_val
  lakeXclim.lt$mei.phi[ii]<-cohXmei$bandp$mn_phs
  lakeXclim.lt$nao.coh[ii]<-cohXnao$bandp$mn_coh
  lakeXclim.lt$nao.p[ii]<-cohXnao$bandp$p_val
  lakeXclim.lt$nao.phi[ii]<-cohXnao$bandp$mn_phs
  lakeXclim.lt$pdo.coh[ii]<-cohXpdo$bandp$mn_coh
  lakeXclim.lt$pdo.p[ii]<-cohXpdo$bandp$p_val
  lakeXclim.lt$pdo.phi[ii]<-cohXpdo$bandp$mn_phs
  lakeXclim.lt$ppt.coh[ii]<-cohXppt$bandp$mn_coh
  lakeXclim.lt$ppt.p[ii]<-cohXppt$bandp$p_val
  lakeXclim.lt$ppt.phi[ii]<-cohXppt$bandp$mn_phs
  lakeXclim.lt$tavg.coh[ii]<-cohXtavg$bandp$mn_coh
  lakeXclim.lt$tavg.p[ii]<-cohXtavg$bandp$p_val
  lakeXclim.lt$tavg.phi[ii]<-cohXtavg$bandp$mn_phs
}

#land, long timescales
NN=length(coherentlakes.lt$lakedata)
ts=c(4,Inf)
for(ii in 1:NN){
  vars<-rownames(coherentlakes.lt$lakedata[[ii]])
  yy<-as.numeric(colnames(coherentlakes.lt$lakedata[[ii]]))
  clnd<-cleandat(coherentlakes.lt$lakedata[[ii]],yy,clev=5)$cdat

  cohXmei<-coh(clnd[vars=="ndvi",], clnd[vars=="mei"], yy, norm="powall", sigmethod="fast")
  cohXnao<-coh(clnd[vars=="ndvi",], clnd[vars=="nao"], yy, norm="powall", sigmethod="fast")
  cohXpdo<-coh(clnd[vars=="ndvi",], clnd[vars=="pdo"], yy, norm="powall", sigmethod="fast")
  cohXppt<-coh(clnd[vars=="ndvi",], clnd[vars=="ppt"], yy, norm="powall", sigmethod="fast")
  cohXtavg<-coh(clnd[vars=="ndvi",], clnd[vars=="tavg"], yy, norm="powall", sigmethod="fast")
  
  cohXmei<-bandtest.coh(cohXmei,ts)
  cohXnao<-bandtest.coh(cohXnao,ts)
  cohXpdo<-bandtest.coh(cohXpdo,ts)
  cohXppt<-bandtest.coh(cohXppt,ts)
  cohXtavg<-bandtest.coh(cohXtavg,ts)
  
  landXclim.lt$mei.coh[ii]<-cohXmei$bandp$mn_coh
  landXclim.lt$mei.p[ii]<-cohXmei$bandp$p_val
  landXclim.lt$mei.phi[ii]<-cohXmei$bandp$mn_phs
  landXclim.lt$nao.coh[ii]<-cohXnao$bandp$mn_coh
  landXclim.lt$nao.p[ii]<-cohXnao$bandp$p_val
  landXclim.lt$nao.phi[ii]<-cohXnao$bandp$mn_phs
  landXclim.lt$pdo.coh[ii]<-cohXpdo$bandp$mn_coh
  landXclim.lt$pdo.p[ii]<-cohXpdo$bandp$p_val
  landXclim.lt$pdo.phi[ii]<-cohXpdo$bandp$mn_phs
  landXclim.lt$ppt.coh[ii]<-cohXppt$bandp$mn_coh
  landXclim.lt$ppt.p[ii]<-cohXppt$bandp$p_val
  landXclim.lt$ppt.phi[ii]<-cohXppt$bandp$mn_phs
  landXclim.lt$tavg.coh[ii]<-cohXtavg$bandp$mn_coh
  landXclim.lt$tavg.p[ii]<-cohXtavg$bandp$p_val
  landXclim.lt$tavg.phi[ii]<-cohXtavg$bandp$mn_phs
}

# test correlation between coherence of lakes and coherence of land
cor.test(lakeXclim.st$mei.coh,landXclim.st$mei.coh)
cor.test(lakeXclim.st$nao.coh,landXclim.st$nao.coh)
cor.test(lakeXclim.st$pdo.coh,landXclim.st$pdo.coh)
cor.test(lakeXclim.st$ppt.coh,landXclim.st$ppt.coh)
cor.test(lakeXclim.st$tavg.coh,landXclim.st$tavg.coh)
# no significant correlations at short timescales

cor.test(lakeXclim.lt$mei.coh,landXclim.lt$mei.coh)
cor.test(lakeXclim.lt$nao.coh,landXclim.lt$nao.coh)
cor.test(lakeXclim.lt$pdo.coh,landXclim.lt$pdo.coh)
cor.test(lakeXclim.lt$ppt.coh,landXclim.lt$ppt.coh)
cor.test(lakeXclim.lt$tavg.coh,landXclim.lt$tavg.coh)
# mostly significant correlations at long timescales

```
Coherence with climate variables was essentially uncorrelated between lakes and the landscape at short timescales, albeit with a small (n=9) number of lakes.
However, at long timescales coherence with climate variables tended to be significantly and positively correlated between lakes and the landscape.
```{r biplots, echo=TRUE, cahce=FALSE}
par(mfrow=c(2,3),mar=c(3.1,3.1,1,1),mgp=c(1.7,0.5,0))
plot(landXclim.st$mei.coh, lakeXclim.st$mei.coh, xlim=c(0,1), ylim=c(0,1), xlab="land", ylab="lake",main="mei")
abline(a=0,b=1)
plot(landXclim.st$nao.coh, lakeXclim.st$nao.coh, xlim=c(0,1), ylim=c(0,1), xlab="land", ylab="lake",main="nao")
abline(a=0,b=1)
plot(landXclim.st$pdo.coh, lakeXclim.st$pdo.coh, xlim=c(0,1), ylim=c(0,1), xlab="land", ylab="lake",main="pdo")
abline(a=0,b=1)
plot(landXclim.st$ppt.coh, lakeXclim.st$ppt.coh, xlim=c(0,1), ylim=c(0,1), xlab="land", ylab="lake",main="ppt")
abline(a=0,b=1)
plot(landXclim.st$tavg.coh, lakeXclim.st$tavg.coh, xlim=c(0,1), ylim=c(0,1), xlab="land", ylab="lake",main="tavg")
abline(a=0,b=1)

par(mfrow=c(2,3),mar=c(3.1,3.1,1,1),mgp=c(1.7,0.5,0))
plot(landXclim.lt$mei.coh, lakeXclim.lt$mei.coh, xlim=c(0,1), ylim=c(0,1), xlab="land", ylab="lake",main="mei")
abline(a=0,b=1)
plot(landXclim.lt$nao.coh, lakeXclim.lt$nao.coh, xlim=c(0,1), ylim=c(0,1), xlab="land", ylab="lake",main="nao")
abline(a=0,b=1)
plot(landXclim.lt$pdo.coh, lakeXclim.lt$pdo.coh, xlim=c(0,1), ylim=c(0,1), xlab="land", ylab="lake",main="pdo")
abline(a=0,b=1)
plot(landXclim.lt$ppt.coh, lakeXclim.lt$ppt.coh, xlim=c(0,1), ylim=c(0,1), xlab="land", ylab="lake",main="ppt")
abline(a=0,b=1)
plot(landXclim.lt$tavg.coh, lakeXclim.lt$tavg.coh, xlim=c(0,1), ylim=c(0,1), xlab="land", ylab="lake",main="tavg")
abline(a=0,b=1)


```

Second, find out how much variability in lakes is explained by coherence with the landscape
```{r wlms lake v landscape, echo=TRUE, cache=TRUE}

## Short timescales------------------------------------------------------------
varexpl.land_v_clim.st<-data.frame(lagoslakeid=names(coherentlakes.st$lakedata),
                                   varexpl.land=rep(NA, length(names(coherentlakes.st$lakedata))),
                                   varexpl.clim=rep(NA, length(names(coherentlakes.st$lakedata))))

#2851 - Walloon Lake
yy<-as.numeric(colnames(coherentlakes.st$lakedata$`2851`))
dat<-list(cleandat(coherentlakes.st$lakedata$`2851`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.st$lakedata$`2851`[2,],yy,clev=5)$cdat)
wlm.lakeland.2851<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.st$varexpl.land[1]<-mean(syncexpl(wlm.lakeland.2851)$syncexpl[syncexpl(wlm.lakeland.2851)$timescales < 4])

#3370 - White Lake
yy<-as.numeric(colnames(coherentlakes.st$lakedata$`3370`))
dat<-list(cleandat(coherentlakes.st$lakedata$`3370`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.st$lakedata$`3370`[2,],yy,clev=5)$cdat)
wlm.lakeland.3370<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.st$varexpl.land[2]<-mean(syncexpl(wlm.lakeland.3370)$syncexpl[syncexpl(wlm.lakeland.3370)$timescales < 4])

#6075 - Fellows Lake
yy<-as.numeric(colnames(coherentlakes.st$lakedata$`6075`))
dat<-list(cleandat(coherentlakes.st$lakedata$`6075`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.st$lakedata$`6075`[2,],yy,clev=5)$cdat)
wlm.lakeland.6075<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.st$varexpl.land[3]<-mean(syncexpl(wlm.lakeland.6075)$syncexpl[syncexpl(wlm.lakeland.6075)$timescales < 4])

#6547 - Lake Moraine
yy<-as.numeric(colnames(coherentlakes.st$lakedata$`6547`))
dat<-list(cleandat(coherentlakes.st$lakedata$`6547`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.st$lakedata$`6547`[2,],yy,clev=5)$cdat)
wlm.lakeland.6547<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.st$varexpl.land[4]<-mean(syncexpl(wlm.lakeland.6547)$syncexpl[syncexpl(wlm.lakeland.6547)$timescales < 4])

#7523 - Annabessacook Lake
yy<-as.numeric(colnames(coherentlakes.st$lakedata$`7523`))
dat<-list(cleandat(coherentlakes.st$lakedata$`7523`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.st$lakedata$`7523`[2,],yy,clev=5)$cdat)
wlm.lakeland.7523<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.st$varexpl.land[5]<-mean(syncexpl(wlm.lakeland.7523)$syncexpl[syncexpl(wlm.lakeland.7523)$timescales < 4])

#28836 - Higginsville Old City Lake
yy<-as.numeric(colnames(coherentlakes.st$lakedata$`28836`))
dat<-list(cleandat(coherentlakes.st$lakedata$`28836`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.st$lakedata$`28836`[2,],yy,clev=5)$cdat)
wlm.lakeland.28836<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.st$varexpl.land[6]<-mean(syncexpl(wlm.lakeland.28836)$syncexpl[syncexpl(wlm.lakeland.28836)$timescales < 4])

#72641 - Lake Wallenpaupack
yy<-as.numeric(colnames(coherentlakes.st$lakedata$`72641`))
dat<-list(cleandat(coherentlakes.st$lakedata$`72641`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.st$lakedata$`72641`[2,],yy,clev=5)$cdat)
wlm.lakeland.72641<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.st$varexpl.land[7]<-mean(syncexpl(wlm.lakeland.72641)$syncexpl[syncexpl(wlm.lakeland.72641)$timescales < 4])

#133500 - Kraut run lake
yy<-as.numeric(colnames(coherentlakes.st$lakedata$`133500`))
dat<-list(cleandat(coherentlakes.st$lakedata$`133500`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.st$lakedata$`133500`[2,],yy,clev=5)$cdat)
wlm.lakeland.133500<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.st$varexpl.land[8]<-mean(syncexpl(wlm.lakeland.133500)$syncexpl[syncexpl(wlm.lakeland.133500)$timescales < 4])

#4909 - Annabelle lake
yy<-as.numeric(colnames(coherentlakes.st$lakedata$`4909`))
dat<-list(cleandat(coherentlakes.st$lakedata$`4909`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.st$lakedata$`4909`[2,],yy,clev=5)$cdat)
wlm.lakeland.4909<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.st$varexpl.land[9]<-mean(syncexpl(wlm.lakeland.4909)$syncexpl[syncexpl(wlm.lakeland.4909)$timescales < 4])

## Long timescales ------------------------------------------------------------
varexpl.land_v_clim.lt<-data.frame(lagoslakeid=names(coherentlakes.lt$lakedata),
                                   varexpl.land=rep(NA, length(names(coherentlakes.lt$lakedata))),
                                   varexpl.clim=rep(NA, length(names(coherentlakes.lt$lakedata))))
# 3370 - White Lake
yy<-as.numeric(colnames(coherentlakes.lt$lakedata$`3370`))
dat<-list(cleandat(coherentlakes.lt$lakedata$`3370`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.lt$lakedata$`3370`[2,],yy,clev=5)$cdat)
wlm.lakeland.3370<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.lt$varexpl.land[1]<-mean(syncexpl(wlm.lakeland.3370)$syncexpl[syncexpl(wlm.lakeland.3370)$timescales > 4])

# 3834 - Hasbrook Lake
yy<-as.numeric(colnames(coherentlakes.lt$lakedata$`3834`))
dat<-list(cleandat(coherentlakes.lt$lakedata$`3834`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.lt$lakedata$`3834`[2,],yy,clev=5)$cdat)
wlm.lakeland.3834<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.lt$varexpl.land[2]<-mean(syncexpl(wlm.lakeland.3834)$syncexpl[syncexpl(wlm.lakeland.3834)$timescales > 4])

# 4243 - unnamed
yy<-as.numeric(colnames(coherentlakes.lt$lakedata$`4243`))
dat<-list(cleandat(coherentlakes.lt$lakedata$`4243`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.lt$lakedata$`4243`[2,],yy,clev=5)$cdat)
wlm.lakeland.4243<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.lt$varexpl.land[3]<-mean(syncexpl(wlm.lakeland.4243)$syncexpl[syncexpl(wlm.lakeland.4243)$timescales > 4])

# 4416 - Green Lake
yy<-as.numeric(colnames(coherentlakes.lt$lakedata$`4416`))
dat<-list(cleandat(coherentlakes.lt$lakedata$`4416`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.lt$lakedata$`4416`[2,],yy,clev=5)$cdat)
wlm.lakeland.4416<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.lt$varexpl.land[4]<-mean(syncexpl(wlm.lakeland.4416)$syncexpl[syncexpl(wlm.lakeland.4416)$timescales > 4])

# 4434 - Kangaroo Lake
yy<-as.numeric(colnames(coherentlakes.lt$lakedata$`4434`))
dat<-list(cleandat(coherentlakes.lt$lakedata$`4434`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.lt$lakedata$`4434`[2,],yy,clev=5)$cdat)
wlm.lakeland.4434<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.lt$varexpl.land[5]<-mean(syncexpl(wlm.lakeland.4434)$syncexpl[syncexpl(wlm.lakeland.4434)$timescales > 4])

# 5895 - Lake Sunapee
yy<-as.numeric(colnames(coherentlakes.lt$lakedata$`5895`))
dat<-list(cleandat(coherentlakes.lt$lakedata$`5895`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.lt$lakedata$`5895`[2,],yy,clev=5)$cdat)
wlm.lakeland.5895<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.lt$varexpl.land[6]<-mean(syncexpl(wlm.lakeland.5895)$syncexpl[syncexpl(wlm.lakeland.5895)$timescales > 4])

# 6199 - unnamed
yy<-as.numeric(colnames(coherentlakes.lt$lakedata$`6199`))
dat<-list(cleandat(coherentlakes.lt$lakedata$`6199`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.lt$lakedata$`6199`[2,],yy,clev=5)$cdat)
wlm.lakeland.6199<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.lt$varexpl.land[7]<-mean(syncexpl(wlm.lakeland.6199)$syncexpl[syncexpl(wlm.lakeland.6199)$timescales > 4])

# 7523 - Annabessacook Lake
yy<-as.numeric(colnames(coherentlakes.lt$lakedata$`7523`))
dat<-list(cleandat(coherentlakes.lt$lakedata$`7523`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.lt$lakedata$`7523`[2,],yy,clev=5)$cdat)
wlm.lakeland.7523<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.lt$varexpl.land[8]<-mean(syncexpl(wlm.lakeland.7523)$syncexpl[syncexpl(wlm.lakeland.7523)$timescales > 4])

# 8369 - unnamed
yy<-as.numeric(colnames(coherentlakes.lt$lakedata$`8369`))
dat<-list(cleandat(coherentlakes.lt$lakedata$`8369`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.lt$lakedata$`8369`[2,],yy,clev=5)$cdat)
wlm.lakeland.8369<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.lt$varexpl.land[9]<-mean(syncexpl(wlm.lakeland.8369)$syncexpl[syncexpl(wlm.lakeland.8369)$timescales > 4])

# 39489 - Ashokan Reservoir West Basin
yy<-as.numeric(colnames(coherentlakes.lt$lakedata$`39489`))
dat<-list(cleandat(coherentlakes.lt$lakedata$`39489`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.lt$lakedata$`39489`[2,],yy,clev=5)$cdat)
wlm.lakeland.39489<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.lt$varexpl.land[10]<-mean(syncexpl(wlm.lakeland.39489)$syncexpl[syncexpl(wlm.lakeland.39489)$timescales > 4])

# 115040 - Ashokan Reservoir East Basin
yy<-as.numeric(colnames(coherentlakes.lt$lakedata$`115040`))
dat<-list(cleandat(coherentlakes.lt$lakedata$`115040`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.lt$lakedata$`115040`[2,],yy,clev=5)$cdat)
wlm.lakeland.115040<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.lt$varexpl.land[11]<-mean(syncexpl(wlm.lakeland.115040)$syncexpl[syncexpl(wlm.lakeland.115040)$timescales > 4])

# 133500 - Kraut Run Lake
yy<-as.numeric(colnames(coherentlakes.lt$lakedata$`133500`))
dat<-list(cleandat(coherentlakes.lt$lakedata$`133500`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.lt$lakedata$`133500`[2,],yy,clev=5)$cdat)
wlm.lakeland.133500<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.lt$varexpl.land[12]<-mean(syncexpl(wlm.lakeland.133500)$syncexpl[syncexpl(wlm.lakeland.133500)$timescales > 4])

# 14815 - unnamed
yy<-as.numeric(colnames(coherentlakes.lt$lakedata$`14815`))
dat<-list(cleandat(coherentlakes.lt$lakedata$`14815`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.lt$lakedata$`14815`[2,],yy,clev=5)$cdat)
wlm.lakeland.14815<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.lt$varexpl.land[13]<-mean(syncexpl(wlm.lakeland.14815)$syncexpl[syncexpl(wlm.lakeland.14815)$timescales > 4])

# 102115 - Wilkinson Lake
yy<-as.numeric(colnames(coherentlakes.lt$lakedata$`102115`))
dat<-list(cleandat(coherentlakes.lt$lakedata$`102115`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.lt$lakedata$`102115`[2,],yy,clev=5)$cdat)
wlm.lakeland.102115<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.lt$varexpl.land[14]<-mean(syncexpl(wlm.lakeland.102115)$syncexpl[syncexpl(wlm.lakeland.102115)$timescales > 4])

# 5463 - Pearl Lake
yy<-as.numeric(colnames(coherentlakes.lt$lakedata$`5463`))
dat<-list(cleandat(coherentlakes.lt$lakedata$`5463`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.lt$lakedata$`5463`[2,],yy,clev=5)$cdat)
wlm.lakeland.5463<-wlm(dat,yy,1,2,norm="powall")
varexpl.land_v_clim.lt$varexpl.land[15]<-mean(syncexpl(wlm.lakeland.5463)$syncexpl[syncexpl(wlm.lakeland.5463)$timescales > 4])

```

Third, find out how much variability in lakes is explained by coherence with variables the landscape is coherent with
```{r wlms lake v climate, echo=TRUE, cache=FALSE}

is.coh_land.lt<-landXclim.lt[colnames(landXclim.lt) %in% 
                               c("lagoslakeid","mei.p","nao.p","pdo.p","ppt.p","tavg.p")]
is.coh_land.lt$mei.p<-ifelse(is.coh_land.lt$mei.p<0.1,1,0)
is.coh_land.lt$nao.p<-ifelse(is.coh_land.lt$nao.p<0.1,1,0)
is.coh_land.lt$pdo.p<-ifelse(is.coh_land.lt$pdo.p<0.1,1,0)
is.coh_land.lt$ppt.p<-ifelse(is.coh_land.lt$ppt.p<0.1,1,0)
is.coh_land.lt$tavg.p<-ifelse(is.coh_land.lt$tavg.p<0.1,1,0)


```


